Measurement of ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross sections in the LArIAT experiment

### 2018-05-25 Chapter-0/0-Introduction.tex
s

\chapter{Introduction}\label{ch:-1}

This thesis work concerns the first measurement of the  ($\pi^-$-Ar)  total hadronic cross section  in the 100-1000 MeV  kinetic energy range and the first measurement of the ($K^+$-Ar) total hadronic cross section  in the 100-650 MeV  kinetic energy range. We performed these measurements at the LArIAT experiment,  a small (0.25 ton)  Liquid Argon Time Projection Chamber (LArTPC) on a beam of charge particles at the Fermilab Test Beam Facility.  Albeit particle and nuclear physics have a long history of hadronic cross section measurements, the work outlined in this thesis presents a new methodology -- the ``thin slice method" -- for cross section measurements possible only thanks to the excellent tracking and calorimetry capabilities of the LArTPC technology. The LArTPC technology allows to see unprecedented details of particle interactions in argon

A renewed interest for precision measurements of hadronic cross sections in argon arises from the current panorama of experimental neutrino physics.

@@

According to the latest Particle Physics Project Prioritization Panel  \cite{P5}, the US particle physics panorama is directing a substantial effort towards the exploration of the Intensity Frontier. In particular, the near future will see the development of a Short Baseline Neutrino Program (SBN) and long baseline neutrino program with  DUNE as far detector. Both these program are based on the Liquid Argon Time Projection Chamber (LArTPC) detector technology. The US liquid argon program has the potential to answer many of the fundamental open questions in particle physics today, such as: is there a fourth generation neutrino? is CP violated in the lepton sector? are there any additional symmetries? And, can we find an indication of Grand Unified Theories?

@@

The long baseline neutrino program and DUNE in particular has a wide physics program. For what it concerns neutrino physics, DUNE has the capability of solving the mass hierarchy and octant problem,  and discovering CP violation in the neutrino sector.
Besides its neutrino program, DUNE can open an experimental window on Grand Unified Theories (GUTs). GUTs could potentially answer fundamental questions such as the existence of non-zero neutrino masses and matter-antimatter asymmetry, explaining some ``accidents" in the Standard Models, such as the exact cancellation of the  proton and the electron charge.  
Probing GUTs at the unification energy scale is impossible by any foreseeable collider experiment. An indirect proof is thus needed such as baryon number violation, which is predicted by almost every GUT in the form of proton decay, bounded nucleon decay or $n-\bar n$ oscillations on long time-scales. Historically, the dominant technology used in these searches has been water Cherenkov detectors, with Super-Kamiokande setting all the current experimental limits on the decay lifetimes at the order of $\sim 10 ^{34}$ years. The DUNE far detector and its non-accelerator physics program is a interesting new actor on this stage.  LArTPCs can in fact complement nucleon decay searches in modes where water Cherenkov detectors are less sensitive, especially $p\rightarrow K^+\bar{\nu}$  \cite{Adams:2013qkq}.

@@

LArIAT \cite{Cavanna:2014iqa}, a small (0.25 ton) LArTPC on a beam of charge particles at the Fermilab Test Beam Facility, is performing precise cross section measurements of charged particles in argon. The LArIAT suite of cross section measurements is a critical set of results which will be employed by both the short- and long- baseline neutrino programs. 
This work is a measurement of the negative pion argon and the positive kaon-argon total hadronic differential cross sections.

@@

In order to measure the kaon-argon differential cross section, several steps were necessary. The analysis starts by identifying a sample of kaons in the beam line and assessing the beam line contaminations. It proceeds with tracking the kaon candidates in the TPC and measuring their calorimetry.  Then,  the hadronic interaction point is identified and distinguished from weak interaction cases. The last step of the analysis, which is the energy unfolding, is now ongoing. 
With  this measurement, I acquired extensive experience in tracking hadrons and measuring their energy inside the TPC. This entailed both an in-depth study of the tracking algorithms performances on data and a dive into the GEANT4 transportation models. Since no previous kaon measurement has been performed on argon, simulation packages such as GEANT4 currently interpolate the kaon transportation on (scarce) data from lighter and heavier nuclei. This is suboptimal for proton decay searches.  In fact, an extremely high signal detection efficiency, measured by testing identification algorithms on simulation, is extremely difficult to obtain if the simulation is not reliable. By adding the experimental data points for kaon-argon cross section to the simulation, my thesis analysis will cover this gap of knowledge.

Chapters:\ref{ch:TheTheory}
Chapter \ref{ch:2}
Chapter \ref{sec:experimentDescription}

>>>>

\chapter*{Introduction}\label{ch:-1}

This thesis work concerns the first measurement of the  ($\pi^-$-Ar)  total hadronic cross section  in the 100-1000 MeV  kinetic energy range and the first measurement of the ($K^+$-Ar) total hadronic cross section  in the 100-650 MeV  kinetic energy range. We performed these measurements at the LArIAT experiment,  a small (0.25 ton)  Liquid Argon Time Projection Chamber (LArTPC) on a beam of charged particles at the Fermilab Test Beam Facility.  Albeit particle and nuclear physics have a long history of hadronic cross section measurements, the work outlined in this thesis presents a new methodology -- the ``thin slice method" -- for cross section measurements, possible only thanks to the detection capabilities of the LArTPC technology. The LArTPC technology allows to see unprecedented details of particle interactions in argon given its full 3D-imaging with millimeter resolution and precise calorimetric reconstruction. A renewed interest for precision measurements of hadronic cross sections, particularly in argon, arises from the current  panorama of experimental  particle physics at the intensity frontier.

@@

According to the latest Particle Physics Project Prioritization Panel  \cite{P5}, the US particle physics panorama is directing a substantial effort towards the exploration of the intensity frontier. In particular, the near future will see the development of a Short Baseline Neutrino Program (SBN) and long baseline neutrino program with  DUNE as far detector. Both these programs are based on the Liquid Argon Time Projection Chamber (LArTPC) detector technology. The US liquid argon program has the potential to answer many of the fundamental open questions in particle physics today, such as: is there a fourth generation neutrino? is CP violated in the lepton sector? are there any additional symmetries? And, can we find an indication of Grand Unified Theories?

@@

DUNE has a vast neutrino and non-accelerator physics reach. For what it concerns neutrino physics, oscillation analyses in DUNE have the capability of solving the mass hierarchy and octant problem,  and discovering CP violation in the neutrino sector. Besides its neutrino program, DUNE can open an experimental window on Grand Unified Theories (GUTs). GUTs could potentially answer fundamental questions such as the existence of non-zero neutrino masses and matter-antimatter asymmetry, explaining some ``accidents" in the Standard Models, such as the exact cancellation of the  proton and the electron charge.   Directly probing GUTs at the unification energy scale is impossible by any foreseeable collider experiment. We then need an indirect proof such as baryon number violation, which is predicted by almost every GUT in the form of proton decay, bounded nucleon decay or $n-\bar n$ oscillations on long time-scales. Historically, the dominant technology used in these searches has been water Cherenkov detectors, with Super-Kamiokande setting all the current experimental limits on the decay lifetimes at the order of $\sim 10 ^{34}$ years. The DUNE far detector and its non-accelerator physics program is a interesting new actor on this stage.  LArTPCs can in fact complement nucleon decay searches in modes where water Cherenkov detectors are less sensitive, especially $p\rightarrow K^+\bar{\nu}$  \cite{Adams:2013qkq}.

@@

LArIAT \cite{Cavanna:2014iqa}, a small (0.25 ton) LArTPC on a beam of charged particles at the Fermilab Test Beam Facility, is performing precise cross section measurements of charged particles in argon. The LArIAT suite of cross section measurements is a critical set of results which will be employed by both the short- and long- baseline neutrino programs. 
This work is a measurement of the negative pion argon and the positive kaon-argon total hadronic differential cross sections.

@@

In order to measure the kaon-argon differential cross section, several steps were necessary. The analysis starts by identifying a sample of kaons in the beam line and assessing the beam line contaminations. It proceeds with tracking the kaon candidates in the TPC and measuring their calorimetry.  Then,  the hadronic interaction point is identified and distinguished from weak interaction cases. The last step of the analysis, which is the energy unfolding, is now ongoing. 
With  this measurement, I acquired extensive experience in tracking hadrons and measuring their energy inside the TPC. This entailed both an in-depth study of the tracking algorithms performances on data and a dive into the GEANT4 transportation models. Since no previous kaon measurement has been performed on argon, simulation packages such as GEANT4 currently interpolate the kaon transportation on (scarce) data from lighter and heavier nuclei. This is suboptimal for proton decay searches.  In fact, an extremely high signal detection efficiency, measured by testing identification algorithms on simulation, is extremely difficult to obtain if the simulation is not reliable. By adding the experimental data points for kaon-argon cross section to the simulation, my thesis analysis will cover this gap of knowledge.\\

This body of work is divided in 8 chapters.
We provide a description of the theoretical framework for the measurements in  Chapter \ref{ch:TheTheory}. Chapter \ref{ch:2} outlines the LArTPC detector technology, while
Chapter \ref{sec:experimentDescription} describes LArIAT experimental setup. We present the event selection for both the pion and kaon analyses, as well as the ``thin-slice method" in Chapter \ref{ch:Interactions}.  Chapter \ref{ch:samples}  describes the work done on the data and Monte Carlo samples in preparation of the cross section analyses.
Chapter \ref{ch:PionXS} shows  the results for the ($\pi^-$-Ar) total hadronic cross section measurement. Chapter \ref{ch:KaonXS} shows  the results for the ($K^+$-Ar) total hadronic cross section measurement. We draw the final remarks on this work in Chapter \ref{ch:Conclusions}

### 2018-05-25 thesis_main.tex
s

\usepackage{geometry} % you need this for yalephd.cls to work.
\usepackage{graphicx}
\usepackage[graphicx]{realboxes}
 % you probably want the rest of these.
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{hyperref}

@@

\setcounter{chapter}{-1}

@@

% Need to define title before the abstract.
\title{Measurement of total hadronic differential cross sections in the LArIAT experiment}
\author{Elena Gramellini}
\advisor{Bonnie T. Fleming}
\date{Date you'll receive your degree} % usually not \today.

@@

\maketitle
\makecopyright{2017}

@@

%Bonnie. I don't even know where to start.
%Flavio
%

%A big thank you to my academic sisters, a bunch of extraordinary women who have taught me so much about myself and our field.
%From the youngest to the oldest (again academically speaking): Supraja, you worked the magic of enriching my past, thank you from the bottom of my heart;  Brooke, the depth of your knowledge, your drive and love for our field has been and will always be a constant inspiration;  Ariana, thanks for keeping me sane for the last four years, if you run for president, remember to delete  our Skype chats, love and hugs to you dudette; and last but not least Xiao, I don't think Fermilab has given me a bigger gift than your friendship.

%Shany
%Jen 
%Jonathan
%Andrzej
%Anto
%LArIAT & MicroBooNE
%Louise

% Starts proper arabic numbering of pages and chapters.
\mainmatter
%\input{Chapters/0-Introduction}
%\input{Chapter-1/1-Theory}
%\input{Chapter-2/2-LArTPCs}
%\input{Chapter-3/3-TheLArIATExperiment}
\input{Chapter-4/4-Methodology}

\input{Chapter-5/5-Preparatory}
\input{Chapter-6/6-PionCrossSection}
\input{Chapter-7/7-KaonCrossSection}

%\input{Chapter-7/7-BackgroundSub}
%\input{Chapter-8/8-CrossSectionAnalysis}\input{Chapter-9/9-Uncertainty}

% Only call appendix once, here.
\appendix
%\input{Chapters/A-KaonAna}
\input{AppendixB-EField/B-LArIATEField}
%\input{Chapters/C-MicroBooNECRT}

>>>>

\usepackage{geometry} % you need this for yalephd.cls to work.
\usepackage{graphicx}
\usepackage[graphicx]{realboxes}
 % you probably want the rest of these.
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{hyperref}
\usepackage[export]{adjustbox}

@@

%\setcounter{chapter}{-1}

@@

% Need to define title before the abstract.
\title{Measurement of ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross sections in the LArIAT experiment}
\author{Elena Gramellini}
\advisor{Bonnie T. Fleming}
\date{Date you'll receive your degree} % usually not \today.

@@

\maketitle
\makecopyright{2018}

@@

I jumped the pond a little less than 5 years ago because I had a taste of what being a physicists in the US might mean and wanted more. 
I came to this vast (and at times inhospitable) land thinking I would learn a lot of physics (which I did) and trying to keep doing research for as long as I could (which I'm fortunate enough to say I'll still do for a while). Little did I know that getting my PhD would be first and foremost a journey to self discovery, the beginning of a path to asymptotically tend to the person and the scientist I want to be. A journey impossible without the help of many people, which I will now proceed to thank, hoping that the acknowledgements won't be longer than the thesis itself.

Mom, you are and will always be the hero of my story; I know this has been challenging for you, thanks for being there for me at every hour of the day and the night, for being always honest with me and for somehow handling this whole US-Italy thing gracefully. I can't believe you still haven't met Bonnie. Ah, and don't worry, I'll always translate things for you.

Bonnie Fleming, my advisor, I don't even know where to start in thanking you. Thanks for listening to me before I proved I was worth be listened to; thanks for always believing I could make it, even at the bottom of my doubts; thanks for the laughters, thanks for leading by example, thanks for  letting me part of the most wonderful research group ;

%Flavio Cavanna 
%Ornella Palamara, I once told Andrzej I think of you as my spirit animal: fierce, independent and direct; he replied I still need to eat a lot of paninis to get there and I can't agree more.

%A big thank you to my academic sisters, from the youngest to the oldest (again academically speaking): Supraja Balasubramanian, you worked the magic of enriching my past, thank you from the bottom of my heart;  Brooke Morrison, the depth of your knowledge, your drive and love for our field has been and will always be a constant inspiration;  Ariana  Hackenburg, thanks for keeping me sane for the last four years, if you run for president -- as you should -- remember to delete our Skype chats, love and hugs to you dudette; and last but not least Xiao Luo, Fermilab has given me lots of extraordinary gifts, but no one  bigger than your friendship.

%Jen Raaf
%Jonathan Asaadi
%Andrzej M. Szelc
%Roberto Acciarri
%LArIAT & MicroBooNE
%To all the members of the NPG, thanks for all the laughters and all the wise insights. You're the smartest bunch of idiots I've ever met, I love y'all.
%Shany %Louise

%To the extraordinary people who stayed close, even with an ocean in between: la Tasso, Marci, la Bea, la Kittirn, Alvin, la Viola e Brogna, Sighi and all the people from the ``La Chiappia Nuova" group, thanks for making me feel like I never left.
%Finally, Anto, cause people might think I'm nice and easy, but you know I'm not. And you're still here.

% Starts proper arabic numbering of pages and chapters.
\mainmatter
\input{Chapter-0/0-Introduction}
\input{Chapter-1/1-Theory}
\input{Chapter-2/2-LArTPCs}
\input{Chapter-3/3-TheLArIATExperiment}
\input{Chapter-4/4-Methodology}
\input{Chapter-5/5-Preparatory}
\input{Chapter-6/6-PionCrossSection}
\input{Chapter-7/7-KaonCrossSection}
\input{Chapter-8/8-Conclusions}

%
%\input{Chapter-9/9-Uncertainty}

% Only call appendix once, here.
\appendix
%%\input{Chapters/A-KaonAna}
\input{AppendixB-TrackingStudies/B-Tracking}
\input{AppendixC-EnergyCalibration/C-EnergyCalibration}
%\input{AppendixB-EField/B-LArIATEField}
%%\input{Chapters/C-MicroBooNECRT}
